Lesson Name:  D_A real data example
Course Name:  Automated Model Selection(AIC,BIC)
Type:         Standard
Author:       Yu Du
Organization: Johns Hopkins University Biostat Department
Version:      2.1.1
================================================================

--- &text
Finally let us look at a real data example with automated model selection.

--- &text
Please read the dataset from http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data directly into R.

--- &cmd_question
Read the data in using `read.csv` and assign the data frame to variable `mpg`

```{r}
mpg <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",sep="")
```
*** .hint
Just type `mpg <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",sep="")` and press Enter.

--- &cmd_question
Change the names of the mpg data frame for easy reference. Make it `"mpg","cyl","displacement","horsepower","weight","acceleration","modelyear","origin","car_name"`.


```{r}
names(mpg) <- c("mpg","cyl","displacement","horsepower","weight","acceleration","model_year","origin","car_name")
```
*** .hint
Just type `names(mpg) <- c("mpg","cyl","displacement","horsepower","weight","acceleration","model_year","origin","car_name")` and press Enter.


--- &cmd_question
Look at the structure of mpg dataset.

```{r}
str(mpg)
```
*** .hint
Just type `str(mpg)` and press Enter.


--- &cmd_question
Look at the summary of mpg dataset.

```{r}
summary(mpg)
```
*** .hint
Just type `summary(mpg)` and press Enter.

--- &text
What we notice here is that horsepower variable is factor variable but what we want is continuous. So let us do the conversion.

--- &cmd_question
Convert the variable horsepower to continuous variable

```{r}
mpg$horsepower <- as.numeric(levels(mpg$horsepower))[mpg$horsepower]
```
*** .hint
Just type `mpg$horsepower <- as.numeric(levels(mpg$horsepower))[mpg$horsepower]` and press Enter.

--- &cmd_question
There are also NA values in the dataset and let us remove all the NA values to have a complete dataset using `complete.cases()`.

```{r}
mpg <- mpg[complete.cases(mpg),]
```
*** .hint
Just type `mpg <- mpg[complete.cases(mpg),]` and press Enter.

--- &cmd_question
Fit a linear model with the formula mpg~cyl+displacement+horsepower+weight+acceleration and assign to `fit`

```{r}
fit <- lm(mpg~cyl+displacement+horsepower+weight+acceleration,data=mpg)
```
*** .hint
Just type `fit <- lm(mpg~cyl+displacement+horsepower+weight+acceleration,data=mpg)` and press Enter.

--- &cmd_question
Do a stepwise regression with backward selection and AIC criterion using `step()` and assign to `st`

```{r}
st <- step(fit)
```
*** .hint
Just type `st <- step(fit)` and press Enter.

--- &cmd_question
Look at a summary of `st`.

```{r}
summary(st)
```
*** .hint
Just type `summary(st)` and press Enter.

--- &text
We can see here that selected model is mpg ~ horsepower + weight, and this model has the best AIC score. This model makes certain sense since common sense is that the higher the horsepower and weight are, the less the mpg is. 

--- &figure

Let us diagnose the best selected model a little bit.

*** .figure
Plot_diagnostic.R

*** .fig_type
new

--- &text
As we can see from the figure, even if this is the best model selected, the diagnosis about the residual does not look quite good. So remember that model selection is not equal to model diagnostic. Also, when we decide whether or not to include a predictor, we need to have a thorough knowledge about the scenario being modelled!


--- &text
This is the end of the class and I do hope you learned something! Thank you!
